Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-22 experiments.tex
draft of comments on results

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
\subsection{Motivation}
In order to evaluate the behavior of the proposed approach
we will be using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The same philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: VLIW architectures let designers
easily trade between power, energy and performances~\cite{}, resulting
in Pareto sets whose extension is an ideal ground for our testing
purposes focused on parameter-space space representation of Paretos.
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
constitutes the key feature to enable the multi-objective trade offs
of the point above. It is statically obtained by the compiler which
performs code trasformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario well suitable for investigating hardly predictable
hardware/software interactions, especially when compared to
architectures that, in order to preserve binary compatibility,
separate software generation processes from the hardware that
will execute it.
\end{itemize}

\subsection{Parameter Space}
System parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiler}. 
The first is directly related to the VLIW core and includes the size
of the register files and the number of functional
units for each type. In particular, five different types of register
files are supported:
GPR (32-bit registers for integers), FPR (64-bit registers for
floating point values) PR (1-bit registers used to store the Boolean
values of predicated instructions), BTR (64-bit registers containing
information about possible future branches) and CR (32-bit control
registers containing information about the internal state of the
processor). The functional units involved are: \emph{Integer units},
\emph{floating point units}, \emph{memory units} (associated with
load/store operations) and \emph{branch units} (associated with branch
operations). With respect to the memory sub-system, the parameters
include \emph{size}, \emph{associativity} and
\emph{block size} for each of the three caches: First-level data cache
(L1D), first-level instruction cache (L1I) and second-level unified
cache (L2U).

@@

As regards the class of benchmark being considered, a set applications representative of some common
frequently running code kernels in an embedded multimedia
environment has been chosen. Table~\ref{tab:bench} shows the set of applications
along with a brief description.
\begin{table}
	\centering
	\caption{Benchmarks}
	\label{tab:bench}
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}
\end{table}

@@

Set of Figures~\ref{fig:pareto_fronts_100}
and~\ref{fig:pareto_fronts_500} show the resultig pareto fronts for two
different scenarios of 100 and 500 simulations budget.

@@

As can be seen TODO...

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a FCP and a VCP exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:dispersion}(a) and (b)
respectively. As it can
be observed, the VCP exploration provides solutions which fall on a
range that is, on average, 23\% and 40\% wider than that provided by a
FCP exploration for power dissipation and execution time,
respectively.
\begin{table}
  \centering
  \begin{tabular}{c}
    \includegraphics[width=0.4\textwidth]{pictures/variation_power.eps} \\
    (a) \\
    \includegraphics[width=0.4\textwidth]{pictures/variation_etime.eps}
    (b) \\
    \includegraphics[width=0.4\textwidth]{pictures/dispersion.eps} \\
    (c) 
  \end{tabular}
  \caption{(a)(b) Variation range for different benchmarks between PS and GA
  and average dispersion absolute dispersion (c)}
  \label{fig:dispersion}
\end{table}

@@

Fig.~\ref{fig:dispersion}(c) shows the average normalised absolute
dispersion errors for different benchmarks for FCP and VCP
exploration. As it can be observed, VCP exploration reduces the
dispersion error on average by 20\% as compared to a FCP exporation.

>>>>

\section{A Case Study: VLIW Hardware/Software Design Space}
\label{sec:ee}
\subsection{Motivation}
In order to evaluate the behavior of the proposed approach
we will be using a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The same philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item{Multi-Objective trade-offs}: VLIW architectures let designers
easily trade between power, energy and performances, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space space representation
\item{Compiler awareness}: Instruction Level Parallelism (ILP)
is statically obtained by the compiler which
performs code trasformations being aware of the underlying hardware
configuration.  This tight hardware/software dependence, makes the
VLIW scenario perfectly suitable for investigating hardly predictable
hardware/software interactions.
\end{itemize}

\subsection{Parameter Space}
In general, system parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiler}. 
The first is directly related to the VLIW core and includes the size
of the register files and the number of functional
units for each type. In particular, five different types of register
files are supported:
GPR (32-bit registers for integers), FPR (64-bit registers for
floating point values) PR (1-bit registers used to store the Boolean
values of predicated instructions), BTR (64-bit registers containing
information about possible future branches) and CR (32-bit control
registers containing information about the internal state of the
processor). The functional units involved are: \emph{Integer units},
\emph{floating point units}, \emph{memory units} (associated with
load/store operations) and \emph{branch units} (associated with branch
operations). With respect to the memory sub-system, the parameters
include \emph{size}, \emph{associativity} and
\emph{block size} for each of the three caches: First-level data cache
(L1D), first-level instruction cache (L1I) and second-level unified
cache (L2U).

@@

As regards the class of benchmark being considered, a set applications representative of some common
frequently running code kernels has been selected. Table~\ref{tab:bench} shows the set of applications
along with a brief description.
\begin{table}
	\centering
	\caption{Benchmarks}
	\label{tab:bench}
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}
\end{table}

@@

Set of Figures~\ref{fig:pareto_fronts_100}
and~\ref{fig:pareto_fronts_500} show the resulting pareto fronts for two
different scenarios of 100 and 500 simulations budget. When possible,
similar algorithm parameters have been adopted. In particular a
population size of 30 has been chosen for GA, corresponding to the
same value for the $K$ of PS. The number of generations (or era)
has been chosen so that the only the given amount of simulations is
actually performed. Thus, when accounting these simulations only unique simulations
has been considered, removing repeations and unfeasibile
configurations.

@@

Low budget Pareto sets are shown in
Figure~\ref{fig:pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly depended upon the application being considered.
These results are obtained in about 5 eras, and the main aim was
demonstrating that with a so limited budget PS still can competer with
genetic approach while even discovering very different Pareto fronts.
Continuing with a more extendent simulation budget as depicted in
Figure~\ref{fig:pareto_fronts_500}, we can qualitatively observe that
even if PS does not stricly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behavior as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc~\ldots.

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective. A
comparison between the variation range for different benchmarks
between a FCP and a VCP exploration for both power dissipation and
execution time is shown in Fig.~\ref{fig:dispersion}(a) and (b)
respectively. As it can be observed, the PS exploration provides solutions which fall on a
range that is, on average, wider than that provided by a
MOGA exploration for power dissipation and execution time, respectively.
\begin{table}
  \centering
  \begin{tabular}{c}
    \includegraphics[width=0.4\textwidth]{pictures/variation_power.eps} \\
    (a) \\
    \includegraphics[width=0.4\textwidth]{pictures/variation_etime.eps} \\
    (b) \\
    \includegraphics[width=0.4\textwidth]{pictures/dispersion.eps} \\
    (c) 
  \end{tabular}
  \caption{(a)(b) Variation range for different benchmarks between PS and GA
  and average dispersion absolute dispersion (c)}
  \label{fig:dispersion}
\end{table}

@@

Fig.~\ref{fig:dispersion}(c) shows the average normalised absolute
dispersion errors for different benchmarks for PS and MOGA
exploration. As it can be observed, MOGA still shows a tendency in
discovering more uniformly dispersed points (i.e. higher values). These results,
together with the variation ranges of Fig.~\ref{fig:dispersion}(a) and
(b) confirm the the availability of density vs extension trade-offs
when comparing PS and multi-objective genetic approaches.

### 2014-07-22 introduction.tex
Results data
draft of comments on results

The design of an embedded system requires several objectives (energy
consumption, area occupation, ...) to be fulfilled. The fulfilling
of these objectives depends on several parameters and in the real
cases no analytical results are known to show the relation between
configurations (combination of parameters) and objectives. Therefore
the simulation is the only way to have a picture of how parameters
impact on the objectives. The problem is that the number of parameters
and the number of values that each of them may take lead to an enormous
number of possible configurations, such that simulating all of them
is an unfeasible activity. Therefore there is the need for the ``Design
Space Exploration'', i.e. a methodology to select a subset of all
possible configurations and simulate only this subset.

@@

The result of every design space exploration is a subset of the configurations
called Pareto Set (\cite{pareto}). Roughly speaking, it is a set of
configurations such that there are no other configurations, among
the simulated ones, that have all the objectives, at the same time,
better fulfilled. If all possible configurations were evaluated, the
ideal Pareto set (the ``true'' Pareto set) could be calculated.
But, as stated above, evaluating all the configurations is not feasable.
Therefore the Pareto set returned by an exploration algorithm is an
approximation of the ``true'' Pareto set.

The ``quality'' of an exploration algorithm involves two aspects:
\begin{itemize}
\item how much the Pareto set returned by the exploration algorithm is similar
to the ``true'' Pareto set (notice that this cannot be evaluated
because the true Pareto set is unknown - it is only an ``ideal''
quality factor)
\item how fast the Pareto set converges towards the final one. If algorithm
$1$ returns the final Pareto set after simulating $n_{1}$ configurations
and algorithm $2$ returns its final Pareto set after $n_{2}$ simulations,
algorithm $1$ is said to be faster than algorithm $2$, iff $n_{1}<n_{2}$.
\end{itemize}

\section{\label{sec:Related-work}Related work}

There are a number of design space exploration algorithms proposed
in literature.

>>>>

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases an high level
system simulation is the only way to have a picture of how system
parameters impact on the objectives.  The key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of parameters, easily yelding an untractable number
of simulations. In order to address such task a ``Design Space
Exploration'' strategy is required, i.e. a methodology to evaluate
only a small subset of all possible configurations while maintaining a
good level of efficiency and accuracy.

The final result of every design space exploration is a subset of the
configurations called Pareto Set (\cite{pareto}). Roughly speaking, it
is a set of configurations representing the trade-offs between
objectives: for each of configuration of the Pareto Set there is no
other configuration performing better for all the objectives
considered. Of course this is an high level estimation so in a
subsequent step of the design flow these few promising configurations
can be more accurately evaluated using detailed low level simulations.

In this work we present PS, a multiobjective design exploration
strategy which introduce the concept of Parameter Space representation
of Pareto Set. The main idea is to ....

\section{\label{sec:Related-work}Related work}
% TODO: short, remove descriptions
% SPEA2 as choice, comparison of it agains other GA is available

@@
